---
title: "ML Assignment Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
# setwd("/users/anchaljaiswal/Downloads")
library(dplyr)
library(ggplot2)
library(car)
library(gridExtra)
library(tidyr)
library(GGally)
getwd()
inp=read.csv("Houses_for_rent_madrid_v5.csv")
df=data.frame(inp)
str(df)
```


## R Markdown

```{r}
mean_price=mean(df$Rent)
df$Outer=factor(df$Outer)
df$Elevator=factor(df$Elevator)
df$Penthouse=factor(df$Penthouse)
df$Cottage=factor(df$Cottage)
df$Duplex=factor(df$Duplex)
df$Semi.detached=factor(df$Semi.detached)

df = df %>%
  mutate(floor_cat=ifelse(Floor<=5,1,ifelse(Floor>=6 & Floor<=10,2,3))) %>%
  mutate(bed_type = ifelse(Bedrooms<=2,1,ifelse(Bedrooms>=3 & Bedrooms<=4,2,3)))
df$floor_cat=factor(df$floor_cat)
df$Area=factor(df$Area)
df$bed_type=factor(df$bed_type)
str(df)

```

```{r}
head(inp)
```


# cleaning data
```{r}
# missing data
sapply(df, function(x) sum(is.na(x)))
```
181 values missing, and on the Floor data

```{r}
# impute cottages as first floor
df[is.na(df$Floor) & df$Cottage == 1,'Floor'] = 1
```



```{r}
# create basement_gf variable: 1 if place is basement or groundfloor, 0 otherwise
table(df$Floor)

df <-
  df %>%
  mutate(basement_gf = ifelse(df$Floor == -1 | df$Floor == 0, 1, 0))

table(df$basement_gf)


```

```{r}
# drop every other row
df <- na.omit(df)
```


```{r}
# check missing values
sapply(df, function(x) sum(is.na(x)))
```

# variable analysis

```{r}
# number of listings by area
ggplot(df,aes(x=reorder(Area,Area,
                     function(x)-length(x)), fill = Region)) +
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Areas")

table(df$Area)
```


```{r}
# how many zones does each area have?
df %>%
  group_by(Area) %>%
  summarise(unique_zones = length(unique(Zone))) %>%
  arrange(desc(unique_zones)) %>%
  ggplot(aes(x = Area, y = unique_zones)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
# rent price distribution by area?
ggplot(df, aes(x = Area, y = Rent, fill = Region)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = mean(df$Rent), color = I('red'), linetype = 2)
```

```{r}
# rent prices by region?
ggplot(df, aes(x = Region, y = Rent, fill = Region)) +
  geom_boxplot() +
  geom_hline(yintercept = mean(df$Rent), color = I('red'), linetype = 2)
```

```{r}
# rent prices for number of bedrooms?
grid.arrange(
  ggplot(df, aes(x = factor(Bedrooms), y = Rent)) +
    geom_boxplot() +
    ggtitle("Average price by bedroom number"),

  ggplot(df,aes(x = Bedrooms)) +
    geom_bar() +
    ggtitle("Number of properties by bedroom number") +
    scale_x_continuous(breaks = seq(0,8,1)),
  ncol = 2
)
```
Price increases in general with more bedrooms, but with larger variance too. One explanation would be that sometimes there are cheap student apartments with large number of bedrooms available, and thus while the property is large, it can hardly be considered a 'valuable' one. 

```{r}
# number of bedrooms by area
df %>%
  group_by(Area) %>%
  ggplot(aes(x = Area, y = Bedrooms, fill = Region)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = mean(df$Bedrooms), color = I('red'), linetype = 2)
```


```{r}
# does having a 'bigger' house increase the rent?
names(df)
ggplot(df, aes(x = Sq..Mt., y = Rent)) +
  geom_point() +
  geom_smooth(method = 'lm', color = I('red'))
```
There seems to be a general pattern, but the lower region has a large variance.
```{r}
ggplot(df, aes(x = Sq..Mt., y = Rent)) +
  geom_point() +
  geom_smooth(method = 'lm', color = I('red')) +
  coord_cartesian()
```




```{r}
# type of house price difference?
grid.arrange(
ggplot(inp, aes(x = factor(Studio), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: studio'),

ggplot(inp, aes(x = factor(Penthouse), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: penthouse'),
ncol = 2
)
```


```{r}
grid.arrange(
ggplot(inp, aes(x = factor(Cottage), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: cottage'),

ggplot(inp, aes(x = factor(Duplex), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: duplex'),

ggplot(inp, aes(x = factor(Semi.detached), y = Rent)) +
  geom_boxplot() +
  ggtitle('Rent: semi detached'),

ncol = 3
)
```



```{r}
ggplot(inp, aes(x = Inside_M30)) +
  geom_bar()

mean_region_m30 <- inp %>%
  group_by(Inside_M30, Region) %>%
  summarise(mean_rent = mean(Rent))

grid.arrange(
ggplot(subset(mean_region_m30, Inside_M30 == 0), aes(x = Region, y = mean_rent, fill = Region)) +
  geom_col() +
  ggtitle("Mean rent regions out of M30"),

ggplot(subset(mean_region_m30, Inside_M30 == 1), aes(x = Region, y = mean_rent, fill = Region)) +
  geom_col() +
  ggtitle("Mean rent regions inside M30"),

ncol = 2
)
```


```{r}
grid.arrange(
  ggplot(subset(inp, Inside_M30 == 0),
         aes(x = Region, y = Rent)) +
    geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
    ggtitle("Rent among Regions out of M30") +
    scale_y_continuous(breaks = seq(0, 17500, 2500), limit = c(0, 17000)),
  
  ggplot(subset(inp, Inside_M30 == 1),
         aes(x = Region, y = Rent)) +
    geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
    ggtitle("Rent among Regions inside M30")+
    scale_y_continuous(breaks = seq(0, 17500, 2500), limit = c(0, 17000)),
  
  ncol = 2
)
```
While the North and South show great variation in rent depending on whether the area is in or out of M30, The South and West, while the rent prices are still higher for areas inside M30, arguably have 'less' variation in Rent prices, whether it is in or out of the M30. In fact, West seems to have some few outliers with higher Rent prices for areas out of M30, indiciating that there may be a affluent suburb in this region out of the busy areas.



```{r}
ggplot(df, aes(x))
```


## Linear Regression

Stepwise method to choose important variables

```{r}
library(MASS)
df_2 <- df %>%
  dplyr::select(-c(Id, Address, bed_type, Zone)) %>%
  na.omit()

null= lm(data = df_2, Rent ~ 1)  # empty model
full = lm(data = df_2, Rent ~ .) # full model

step = stepAIC(null, scope=list(lower=null, upper=full), direction = "forward", na.rm = T)
?stepAIC
step$anova
summary(step)
```




```{r pressure, echo=FALSE}
fit2= lm(data=df, Rent ~ Sq..Mt. + floor_cat + Penthouse + Cottage  + Region + Outer)
summary(fit2)
```


```{r}
fit3= lm(data=df, Rent ~ Sq..Mt. + floor_cat + Cottage + Region + Duplex + Inside_M30)
summary(fit3)
```


```{r}

fit4 = lm(data=df_2, Rent ~ Sq..Mt. + Area + Inside_M30 + Outer + Penthouse + Cottage + Studio + Bedrooms + Duplex)

summary(fit4)
anova(fit4)
#R2 = 0.7589

fit5 = lm (data=df_2, Rent ~ Sq..Mt. + Area + Inside_M30 + Outer + Penthouse + Cottage + Bedrooms)
summary (fit5)
anova(fit5)
#R2 = 0.7577

fit6 = lm (data = df_2, Rent ~ Sq..Mt. + Area + factor(Inside_M30) + Outer + Penthouse)
summary(fit6)
anova(fit6)
#R2 = 0.7574 

#So the best model (all IV significant) is model fit6

```

```{r}
ggplot(df, aes(x = Rent)) +
  geom_histogram(color = I('white')) +
  ggtitle("Distribution of rent") +
  geom_vline(xintercept = mean(df$Rent), color = I('red'), linetype = 2) +
  geom_vline(xintercept = median(df$Rent), color = I('blue'), linetype = 2)
```
Outliers pulling the rent price up, but the median price is around 1400, while the mean is a little higher at 1946. Around 50% of the prices are between 2500 and 950.
```{r}
summary(df$Rent)
```


```{r}
table(df$Area)
```


## VERIFY STATISTICAL ASSUMPTIONS OF FIT 6

```{r}

#heterocedasticity

qplot(predict(fit6), rstandard(fit6)) + geom_hline(yintercept = 0) + geom_hline(yintercept = 2, colour=I("red"), alpha=I(0.05)) + geom_hline(yintercept = -2, colour=I("red"), alpha=I(0.05))


qplot(predict(fit6), rstandard(fit6)) + geom_hline(yintercept = 0) + geom_hline(yintercept = 2, colour=I("red"), alpha=I(0.05)) + geom_hline(yintercept = -2, colour=I("red"), alpha=I(0.05))

#normality of residuals

q1 = qplot(rstandard(fit6), geom="blank") + geom_histogram(aes(y=..density..), colour=I("gray"), binwidth=0.5)+ stat_function(fun=dnorm, args=list(mean=0, sd=1), colour=I("red"), alpha=I(0.5))

q2 = qplot(sample=rstandard(fit6)) +
  geom_abline(slope=1,intercept=0)

grid.arrange(q1, q2, nrow=1)


#independence of residuals

durbinWatsonTest(fit6)


#Global Linear Model Assumptions

gvlma(fit6)


```

So, apparently our model is heterocedastic, residuals are almost independent (Durbin-Watson = 1.77, when it should be 2) and somehow follow a Normal Distribution. It is not a perfect one probably due to the impact of outliers. 




## REGRESSION DIAGNOSTIC

```{r}
#Multicollinearity

vif(fit6)

#Result is not > 5 for any variable so there is no collinearity.

#Outliers

outlierTest(fit6)

#There are significant outliers.


```


```{r}
ggplot(df, aes(x = factor(Cottage), y = Rent)) +
  geom_boxplot()
```

