---
title: "segmentation_analysis"
output:
  html_document: default
  html_notebook: default
---

```{r}
set.seed(1680)

library(dplyr)
library(ISLR)
library(cluster)
library(Rtsne)
library(ggplot2)
library(gridExtra)
```



```{r}
getwd()
list.files()

rc <- read.csv("raw_trips_final.csv")
```


```{r}
# look at structure
str(rc)

# head
head(rc)

names(rc)
```


```{r}
summary(rc)
```


```{r}
# clean data
rc$dep_month <- factor(rc$dep_month)
rc$dep_dow <- factor(rc$dep_dow)
rc$insurance <- factor(rc$insurance)

```

```{r}
head(rc)
```

```{r}
# drop unneeded variables
rc <- rc %>%
  select(-departuredate, -returndate, -purchasedate, -seats)
```


```{r}
rc_vars <- rc %>%
  select(id, Z_age, travel_companion, weekend_trip)
```


# calculate distance
Defining (dis)similarity between observations
Euclidean is only for continuous variables, and is not applicable for mixed variables.
Gower distance can handle mixed data types.
Gower distance: for each var type, a particular metric that works well for that type is used and scaled to fall between 0 and 1 (e.x. - Manhattan for quantiative intervals, rank & manhattan for ordinal, dice coefficietn for nominal a.k.a., categorical vars)


```{r}
gower_dist_rc <- daisy(rc_vars[,-1],
                       metric = "gower")

summary(gower_dist_rc)
```

```{r}
gower_mat_rc <- as.matrix(gower_dist_rc)
```

```{r}
# find best number of k clusters
sil_width_rc <- c(NA)
for(i in 2:8) {
  pam_fit <- pam(gower_mat_rc,
                 diss = T,
                 k = i)
  sil_width_rc[i] <- pam_fit$silinfo$avg.width
}

# plot silhouette
qplot(x = c(1:8), y = sil_width_rc, geom = "point") +
  geom_line()

```

```{r}
# max silhoutte score and index
max(sil_width_rc, na.rm = T)
which.max(sil_width_rc)
```


```{r}
pam_fit_rc <- pam(gower_dist_rc, diss = T,
                  k = 4)

pam_results_rc <- rc_vars %>%
  select(-id) %>%
  mutate(cluster = pam_fit_rc$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results_rc$the_summary
```

```{r}
# find out what values are the medoids
rc[pam_fit_rc$medoids,]
```



```{r}
# create lower dimensional space to visualize many variables at the same time
# here, t-SNE (t-distributed stochastic neighborhood embedding) is used
tsne_obj_rc <- Rtsne(gower_dist_rc, is_distance = T)

tsne_data_rc <- tsne_obj_rc$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit_rc$clustering),
         name = rc_vars$id)

ggplot(aes(x = X, y = Y), data = tsne_data_rc) +
  geom_point(aes(color = cluster))
```


```{r}
# add the cluster number for each row back to the original dataset
clu <- tsne_data_rc[,'cluster']

rc <- rc %>%
  mutate(cluster = clu)
```


```{r}
head(rc, 3)
```

```{r}
# rename each cluster by unique name
rc$cluster <- as.character(rc$cluster)
rc$cluster[rc$cluster == 1] <- "big_family_retreat"
rc$cluster[rc$cluster == 2] <- "comfort_duo"
rc$cluster[rc$cluster == 3] <- "young_family"
rc$cluster[rc$cluster == 4] <- "young_and_free"
rc$cluster <- as.factor(rc$cluster)
```

```{r}
head(rc,4)
```


```{r}
rc %>%
  group_by(cluster) %>%
  summarise(count = n(),
            pert = count / dim(rc)[1]) 
```


# comparing with profiling variables
```{r}
ggplot(rc, aes(cluster, fill = travel_companion)) +
  geom_bar() +
  ggtitle("Travel companion by cluster")
```

```{r}
ggplot(rc) +
  geom_freqpoly(aes(age,..density.., color = cluster)) +
  geom_freqpoly(aes(age, ..density..), linetype = 2) + 
  ggtitle("Density of age distribution by cluster")
```



```{r}
ggplot(rc,aes(cluster, age)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
  stat_summary(aes(label=sprintf("%0.2f",round(..y..,digits = 2))), 
               fun.y=mean, geom="text", size=2, vjust = -1.5) +
  ggtitle("Distribution of age by cluster") +
  geom_hline(yintercept = mean(rc$age), color = I('blue'), linetype = 2)
```

```{r}
rc %>%
  group_by(cluster) %>%
  summarise(median = median(age))
```

```{r}
names(rc)
```


```{r}
  ggplot(rc,aes(x= standard_holiday, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Proportion of people going to standard holiday by cluster")

```


```{r}

ggplot(subset(rc, standard_holiday == "True"), aes(x = standard_holiday, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 2) +
  ggtitle("Number of people going on standard holiday months by cluster")
```


```{r}
ggplot(rc) +
  geom_freqpoly(aes(duration,..density.., color = cluster)) +
  geom_freqpoly(aes(duration, ..density..), linetype = 2) + 
  ggtitle("Density of trip duration distribution by cluster")
```


```{r}
ggplot(rc,aes(cluster, duration)) +
  geom_boxplot() + 
  stat_summary(fun.y = mean, geom = 'point', color = I('red')) +
  stat_summary(aes(label=sprintf("%0.2f",round(..y..,digits = 2))), 
               fun.y=mean, geom="text", size=2, vjust = -1.5) +
  ggtitle("Duration of trip by cluster") +
  geom_hline(yintercept = mean(rc$duration), color = I('blue'),
             linetype = 2)
```

```{r}
  ggplot(rc,aes(x= weekend_trip, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Proportion of people going on weekend trip by cluster")

```

```{r}
ggplot(subset(rc, weekend_trip == "True"), aes(x = weekend_trip, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 2) +
  ggtitle("Number of people going on weekend trip by cluster")
```

```{r}
  ggplot(rc,aes(x= gender, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Gender proportion by cluster")
```



```{r}
ggplot(rc) +
  geom_bar(aes(cluster, fill = gender), position = 'dodge') +
  ggtitle("Gender count by cluster")
```

```{r}
  ggplot(rc,aes(x= children, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Proportion of people accompanying children by cluster")
```

```{r}
ggplot(subset(rc, children == "True"), aes(x = children, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 2) +
  ggtitle("Number of people accompanying children by cluster")
```


```{r}
ggplot(rc) +
  geom_freqpoly(aes(purchase_lag,..density.., color = cluster)) +
  geom_freqpoly(aes(purchase_lag, ..density..), linetype = 2) +
  ggtitle("Density of purchase lag distribution by cluster")

```

```{r}
ggplot(rc, aes(x = cluster, y =purchase_lag)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, color = I('red'), geom = 'point') +
  stat_summary(aes(label=sprintf("%0.2f",round(..y..,digits = 2))), 
               fun.y=mean, geom="text", size=2, vjust = -1.5) +
  geom_hline(yintercept = mean(rc$purchase_lag), color = I('blue'), linetype = 2) + 
  ggtitle("Distribution of purchase lag by cluster")
```


```{r}
ggplot(rc,aes(x= customer_type, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Proportion of people for each customer type by cluster")
```


```{r}
ggplot(rc) +
  geom_bar(aes(cluster, fill = customer_type), position = 'dodge') +
  ggtitle("Number of people for each customer type by cluster")
```

```{r}
ggplot(rc,aes(x= car, y=..prop..,group = 1,fill = cluster)) + 
    stat_count(show.legend = F) + 
    facet_wrap(~cluster) +
    geom_text(stat = 'count', 
              aes(label = sprintf("%0.2f",
                                  round(..prop.., digits = 2))),
              vjust = -0.5, size = 1) +
  ylab("Proportion") +
  ggtitle("Proportion of people renting car by cluster")
```

```{r}
ggplot(subset(rc, car == "True"), aes(x = car, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count',
            aes(label = ..count..),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 2) +
  ggtitle("Number of people renting car by cluster")
```



```{r}
ggplot(rc, aes(x=hotel, y=..prop..,group = 1,fill = cluster)) +
  stat_count(show.legend = F) + 
  facet_wrap(~cluster) +
  geom_text(stat = 'count', 
            aes(label = sprintf("%0.2f",
                                round(..prop.., digits = 2))),
            size = 2,
            vjust = -0.5) +
  ylab("Proportion") +
  ggtitle("Proportion of people renting hotel by cluster")
```

```{r}
ggplot(subset(rc, hotel == 'True'),
       aes(x = hotel, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count', 
            aes(label = ..count..),
            size = 2,
            position = position_dodge(width = 0.9),
            vjust = -0.5) +
  ggtitle("Number of people renting hotel by cluster")
```



```{r}
ggplot(rc, aes(x=insurance, y=..prop..,group = 1,fill = cluster)) +
  stat_count(show.legend = F) + 
  facet_wrap(~cluster) +
  geom_text(stat = 'count', 
            aes(label = sprintf("%0.2f",
                                round(..prop.., digits = 2))),
            size = 2,
            vjust = -0.5) +
  ylab("Proportion") +
  ggtitle("Proportion of people with insurance by cluster")
```

```{r}
ggplot(subset(rc, insurance == 'True'),
       aes(x = insurance, fill = cluster)) +
  geom_bar(position = 'dodge') +
  geom_text(stat = 'count', 
            aes(label = ..count..),
            size = 2,
            position = position_dodge(width = 0.9),
            vjust = -0.5) +
  ggtitle("Number of people with insurance by cluster")
```


```{r}
grid.arrange(
rc %>%
  filter(cluster == "big_family_retreat") %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))) +
  ylab(""),

rc %>%
  filter(cluster == "big_family_retreat") %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))) +
  ylab(""),
ncol = 2,
top = "Departure month and day of week proportion for 'Big family retreat' cluster")
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == "comfort_duo") %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))) +
  ylab(""),

rc %>%
  filter(cluster == "comfort_duo") %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))) +
  ylab(""),
ncol = 2,
top = "Departure month and day of week proportion for 'Comfort duo' cluster")
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == "young_and_free") %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))) +
  ylab(""),

rc %>%
  filter(cluster == "young_and_free") %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..)))+
  ylab(""),
ncol = 2,
top = "Departure month and day of week proportion for 'Young and free' cluster")
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == "young_family") %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))) +
  ylab(""),

rc %>%
  filter(cluster == "young_family") %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))) +
  ylab(""),
ncol = 2,
top = "Departure month and day of week proportion for 'Young family' cluster")
```



```{r}
write.csv(rc, "raw_trips_final.csv")
```

